---
title: "04 Smoking 2"
tags: [Notebooks/Thinking About Data]
output: 
  html_notebook: 
    toc: yes
---

# Comparison of Samples with *Student*'s $t$ test
## Preamble

```{r}
# Preamble
## Install Pacman
load.pac <- function() {
  
  if(require("pacman")){
    library(pacman)
  }else{
    install.packages("pacman")
    library(pacman)
  }
  
  pacman::p_load(xts, sp, gstat, ggplot2, rmarkdown, reshape2, ggmap,
                 parallel, dplyr, plotly, tidyverse, reticulate, UsingR, Rmpfr,
                 swirl, corrplot, gridExtra, mise, latex2exp, tree, rpart, lattice)
  
}

load.pac()
mise()
load(file = "./TAD.rdata")
```

Load the Data

```{r}
(birthwt <- as_tibble(read.csv(file = "./0datasets/birthwt.csv", header = TRUE, sep = ",")))

 # Always use doubles / numeric unless the data is legitamately
   # restricted to the integers, it will break functions ( and so it should)
birthwt$bwt <- as.numeric(birthwt$bwt)

birthwt$smoke <- c(FALSE, TRUE)[birthwt$smoke]

summary(birthwt)
str(birthwt)
dim(birthwt)
nrow(birthwt)
ncol(birthwt)
```


## Smoking Data
### Permutation Test
Previously the birth weights from smoking where compared by creating a
hypothesis test:

+ $H_0: \quad$ The two Populations have mean values that are equal
+ $H_a: \quad$ The two Populations have mean values that are equal



1. Simulating the data 
### Wilcoxon-Mann-Whitney Test
### Standardisation
### Permutation Again
### Confidence Intervals
### Bootstrap
### Using the **_R_** Library `boot`
### *Student*'s $t$ Distribution

## Sales Data
## Spider Data
